---
title: "Test Scores Report"
output: pdf_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(knitr)
data <- read_csv("https://raw.githubusercontent.com/stenhaug/sdp_workshop/master/ECLS_Sample_1.csv")
```

# Understanding the data

There are `r nrow(data)` students in the data. These students have a total of `r n_distinct(data$teach_id)` unique teachers and `r n_distinct(data$school_id)` unique schools.

Let's take a peek at the data:

```{r}
kable(head(data))
```

Let's see how many times each of the top 10 most common schools show up in the data

```{r, echo = TRUE}
data %>% 
  count(school_id, sort = TRUE) %>% 
  slice(1:10) %>% 
  kable()
```

# Relationship between lit and math pretests

```{r, warning = FALSE, message = FALSE}
data %>% 
  ggplot(aes(x = lit_pretest, y = math_pretest)) +
  geom_point() +
  geom_smooth() +
  labs(x = "Lit Pretest", y = "Math Pretest", title = "Positive relationship between pretests")
```

# Growth

## Histograms of growth

```{r, warning = FALSE, message = FALSE}
hist(data$lit_posttest / data$lit_pretest, xlab = "Multiplier", ylab = "Count", main = "Lit Growth")
```

```{r, warning = FALSE, message = FALSE}
hist(data$math_posttest / data$math_pretest, xlab = "Multiplier", ylab = "Count", main = "Math Growth")
```

## Relationship between growth

```{r, warning = FALSE, message = FALSE}
data_with_growth <- data %>% 
  mutate(lit_growth = lit_posttest / lit_pretest,
         math_growth = math_posttest / math_pretest) 

data_with_growth %>% 
  ggplot(aes(x = lit_growth, y = math_growth)) +
  geom_point() +
  geom_smooth(method = "lm")
```

## Predicting math growth based on covariates

```{r}
model_math_growth <- lm(math_growth ~ ses + female + black + hisp + asian, data = data_with_growth)

summary(model_math_growth)
```

For example, the model finds that the coefficient on SES is `r round(summary(model_math_growth)$coef[2, 1], 2)`. This tells us that...

# Ideas for things to add

- Model lit growth as well
- Only some covariates were included, can you include other covariates?
- Can you turn the insights from the model into crisp visualizations?

